---
title: "CUNY MSDS Data 607 HW #4"
author: "Samantha Deokinanan"
date: "3rd March, 2019"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE  
    toc_depth: 4
---
***
### Overview {.tabset .tabset-fade .tabset.-pills}
Traveling by air is, contrary to popular opinion, the safest way to travel. Any incident with the airplane itself usually occur before take-off or when the airplane is landing. Nonetheless, this still leads to hundreds of frustrated flyers due to delays. 

In this assignment, a data set of *ALASKA* and *AM WEST* on-time and delayed number of flights is giving for five cites. With this information, there will be an investigation to answer the following questions:

* Which airplane had a better flight time for each destination?
* What is the probability of seeing fewer delays for each destination?

***
#### The Packages

This assignment was accomplished by utilizing these packages for both data analysis and visualizations.
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library("tidyr")
library("dplyr")
library("kableExtra")
library("ggplot2")
```
***
#### The Data

The data was recorded in a .csv file and imported into R via [my GitHub](https://github.com/greeneyefirefly/Data607/tree/master/HomeWork/HW4) (where you can also find the Rmd for this page). You will see below that the data is not in a very clean form to conduct analysis easily, therefore this data set needed to be tidy.

```{r}
theURL <- "https://raw.githubusercontent.com/greeneyefirefly/Data607/master/HomeWork/HW4/TTD.csv"
untidyraw <- data.frame(read.csv(file = theURL, header = TRUE, sep = ","))
```
***
### Data Manipulation {.tabset .tabset-fade .tabset.-pills}
A couple data manipulations were required before the data set was presentable and suitable for use.

* Visual Cleaning - making the data frame variables comprehensible 
```{r}
# Remove the NAs
untidy <- untidyraw[-c(3),]
# Identify the airplanes for each flight status
untidy[c(2,4),1] <- untidy[c(1,3),1]
# Column and row renaming
colnames(untidy)[c(1,2)] <- c('Airline','Status')
rownames(untidy) <- 1:nrow(untidy)
```

* Tidying and Transforming - tidying for coherent analyses. To answer the proposed questions, the data set was transformed to a better form so that flights were separated by their status.

```{r}
# Gather the data by city, then spread it by flight status.
tidy <- spread(gather(untidy, "Destination", "Time", 3:7), "Status", "Time")
# Convert character data into factor
tidy <- mutate_if(tidy, is.character, as.factor)
tidy$Airline<-factor(tidy$Airline)
```
***
Here is the untidy, raw data set and what is looks like after the tidying and transformations.

#### Untidy
```{r echo=FALSE}
kable(untidyraw) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```
***
#### Tidy
```{r echo=FALSE}
kable(tidy) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```
***
##### Calculations
At this step, any new variables needed to start the analysis were determined, and added to the tidy flight data frame. These include:
```{r}
# Total number of flights
tidy$TotalFlights <- rowSums(tidy[,c(3, 4)])
# Probability of on-time flight status
tidy$ProbOnTime <- round((tidy$`on time`/ (tidy$TotalFlights)), digits = 3)
# Probability of delayed flight status
tidy$ProbDelay <- round((tidy$delayed/ (tidy$TotalFlights)), digits = 3)
```
***
### Analysis & Results {.tabset .tabset-fade .tabset.-pills}

#### The Better Flight
If an airplane had more on-time status than delayed, that airplane had a better flight time status. The result suggested that the flight with the lowest probability of delayed flights was *ALASKA*, with an average delayed probability of 11.2%. 

```{r}
# Average probability of delayed flights
aggregate(ProbDelay ~ Airline, tidy, mean)
```

Moreover, Figure 1 further depicts the difference in delayed flights for each destination by comparing *ALASKA* to *AM WEST*. When it pertains to overall flight status, both airplanes have more delayed flights when travelling to San Francisco while both airplanes have less delayed flights when travelling to Phoenix. Nonetheless, *ALASKA* still had less delayed flights than *AM WEST* for each destination.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Probability of Delayed Flights by Airline for Each City
a <- aggregate(ProbDelay ~ Airline + Destination, tidy, mean)
ggplot(a, aes(x = Destination , y = a$ProbDelay, colour = Airline)) + geom_point(size = 4, aes(shape = factor(Airline))) + geom_text(aes(label = a$ProbDelay), vjust = 1.75, hjust = .5, position = position_dodge(width = 2), color = "black") + theme(plot.background = element_rect(fill = 'light blue')) + ggtitle("Figure 1. Probability of Delayed Flights by Airline for Each City") + ylab("Probability of Delayed Flights") + coord_flip()
```

***
#### Predictor of Delays

What made flights to Phoenix less likely to be delayed than San Francisco? Even though there are numerous variables that can be accountable to cause this, such as weather, the economy, etc., using the count data available, it can investigated if the expected counts of delayed flights are influenced by the destination based on the Airline flight records. 

A Poisson regression model was designed to investigate the number of delayed flights based on the airline, destination, and total flights as predictors. The response variable is the number of `delayed` recorded at each of five different `Destination` and by the two `Airline` categories. The difference in the total number of flights will be accounted for with the log exposure variable `TotalFlights`.

The results below shows that the model fits reasonably well because the goodness-of-fit chi-squared test is not statistically significant , p > 0.05. The indicator variable `Destination` is compared among four cities with `Los Angeles` being the reference category, the expected log count for `San Diego` was the only non-significant indicator variable at 95% CI. Specifically, the expected log count for `Phoenix` is 0.66 lower than the expected log count for `Los Angeles`. The expected log count for `San Francisco` is 0.59 higher than the reference. To determine if `Destination` itself, overall, is statistically significant, the model was compare to another without `Destination`. 

```{r}
# Poisson Regression of delayed flights on the predictors
model <- glm(delayed ~ offset(log(TotalFlights)) + Destination + Airline, family=poisson(link=log), data = tidy)
summary(model)
with(model, cbind(res.deviance = deviance, df = df.residual,
  p = pchisq(deviance, df.residual, lower.tail=FALSE)))
```

Here, a test of the overall effect of `Destination` is done by comparing the deviance of the full model with the deviance of the model excluding `Destination`. The four degree-of-freedom chi-square test indicates that `Destination`, taken together, is a statistically significant predictor of `delayed`.

```{r}
# Update Model 1 by dropping the destination
model2 <- update(model, . ~ . - Destination)
# Test the model differences with Chi-Square test
anova(model2, model, test="Chisq")
```

***
#### I Choose You!!!
Lastly, a graph of the `Before` probability of delayed flights and `After` probability of the predicted number of delayed flight indicates that the most delayed flights are still predicted for *AM WEST* for all the destination if the travel the some number of times. Specifically, while the actual count data showed *ALASKA* had a better flight rate for `Seattle`, `San Francisco` and `San Diego`, the prediction suggests there will likely be a bit more delays than compared to the same flights with *AM WEST*, which are less. Moreover, *AM WEST* is expected to still have similar delays when flying to `Phoenix` while *ALASKA* has a better chance of being on-time, this is similar for flights to `Los Angeles`.

```{r echo=FALSE, warning=FALSE}
predict <- predict(model, type="response")
ratio <- data.frame(tidy$Airline, tidy$Destination, predict, tidy$TotalFlights, tidy$ProbDelay)
ratio$ProbPre <- round(predict/tidy$TotalFlights, 3)
ratio <- gather(ratio, "tidy.ProbDelayed", "ProbPre", 5:6)
ratio <- mutate_if(ratio[-c(3,4)], is.character, as.factor)
colnames(ratio) <- c('Airline','Destination', 'Status', 'Probability')
levels(ratio$Status) <- c("After","Before")

# Creating a plot of the predicted values
ggplot(ratio, aes(x = Destination, y = Probability, color = Status)) + geom_point(aes(fill = Airline, shape = factor(Airline)), size = 3) + facet_grid(. ~ Airline) + theme(plot.background = element_rect(fill = 'light blue')) + ggtitle("Figure 2. Predicted Probability of Delayed Flights by Airline for Each City") + ylab("Probability of Delays") + coord_flip()
```

***
### Conclusions
In conclusion, while both airplanes have more delayed flights when travelling to San Francisco and less delayed flights when travelling to Phoenix, *ALASKA* has less delayed flights than *AM WEST* for each destination. This could be, as the Poisson regression suggests, that flight destination is a statistically significant predictor of delayed flights. When this prediction was further analyzed, it is predicted that *ALASKA* will eventually see more delayed flights while *AM WEST* may not for `Seattle`, `San Francisco` and `San Diego` destinations, than the other two destinations, but the probability of delayed flights is still higher for *AM WEST* overall.

***
### Works Cited
Tinsley, Howard E. Handbook of Applied Multivariate Statistics and Mathematical Modeling. Academic Press, 2006.